---
title: "R Markdown Basics"
author: "Shilaan Alzahawi"
date: "`r Sys.Date()`"
output: html_document
---

# Rendering a document

* Knit button  
* Windows: ctrl + shift + k 
* Mac: command + shift + k 

# Formatting Basics 

Help > Markdown Quick Reference  

## Emphasis

*italic* **bold**
_italic_ __bold__

## Headers

# Header 1
## Header 2
### Header 3 
#### Header 4 

## Manual Line Breaks 

End a line with two or more spaces.

## LaTeX Equations

R Markdown supports $\LaTeX$ syntax. For example:

We ran a two-sample *t*-test with $\alpha = .05$ and $\beta = .10$. 

## Tables

First Header  | Second Header
------------- | -------------
Content Cell  | Content Cell
Content Cell  | Content Cell

# Code basics

## Code chunks

Insert code chunks:  
* Insert button  
* Windows: control + alt + i  
* Mac: command + option + i   

```{r}
data(sleep)
```

### Chunk options
It's recommended to name your code chunks (for instance, below, we label the chunk "get-info"). If you need a separator, hyphens (-) are recommended. Avoid spaces, periods (.), and underscores (_). 

The behavior of a chunk can be controlled by passing additional comma-separated arguments.  

* echo = FALSE removes the code from the rendered document  
* warning = FALSE removes warning messages from the rendered document
* include = FALSE removes chunk output from the rendered document (nothing is written in the output document, but the code is still evaluated so it can be used later)

```{r, mean-sleep, include = FALSE}
mean(sleep$extra)
```

## Inline code

To avoid copy-pasting, we can execute R code in-line using backticks. For instance,  

M = `r mean(sleep$extra)` should yield M = 1.54.

# Citations

If you don't have a reference manager yet, I recommend Zotero for its convenient integration with citr.  

* Download Zotero at https://www.zotero.org  
* Download the Better BibTex for Zotero extension https://retorque.re/zotero-better-bibtex/   
* Install the citr package  

```{r install-citr, eval = FALSE}
devtools::install_github("crsh/citr")
```


| **Citation Type** | **Syntax** | **Rendered Citation** |
| ------------- | ------------- | ------------- | 
| Citation within parentheses | [@dienes2016] | (Dienes, 2016) |
| Multiple citations  | [@gigerenzer2004; @goodman2019]  | (Gigerenzer, 2004; Goodman, 2019)  | 
| In-text citations | @lakens2014 | Lakens (2014) |
| Year only | [-@oakes1986] | (1986) |
| Prefix | [e.g., @wagenmakers2007] | (e.g., Wagenmakers, 2007)
| Suffix | [@neyman1957: p. 10] | (Neyman, 1957: p. 10)

# Installing papaja

First, we install TinyTex (this allows us to render PDF documents).   

```{r, install-tex, eval = FALSE}
install.packages("tinytex") 
tinytex::install_tinytex() 
```

Second, we install papaja.  

```{r, install-papaja, eval = FALSE}
install.packages("devtools") 
devtools::install_github("crsh/papaja@devel") 
```


# Acknowledgement

Part of today's tutorial is based on materials generously provided by Frederik Aust at https://osf.io/r5h4m/

# References

Frederik Aust & Marius Barth. (2019). An alternative to copy-paste reporting: Create dynamic, submission-ready, APA-style manuscripts in R with papaja. *European Conference on Visual Perception.* https://osf.io/r5h4m/